An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-12-29 EconGrowth2.Rnw
Edited description of resample to include reindexing.

Under that assumption, it would not be surprising if the residual from 
any one year had been observed in a different year. 
By applying randomly selected residuals (with randomly selected signs)
to model predictions at each year of the GDP time series, we create a resampled
GDP time series ($y_{resample,t}$) that can be fitted using the techniques
described below.

\begin{equation} \label{eq:resample_y}
  y_{resample,t} = y_{pred,t} + \mathrm{rand}(-1, 1) \cdot \mathrm{abs}(\mathrm{rand}(r_t))
\end{equation}

We repeated the resampling process \Sexpr{numResamples()} times for each
combination of growth model, energy quantification, and country.

@@

It is important to note that large residuals can arise from 
(a) poor quality historical time series data ($y_{act}$) or 
(b) a mathematical model that does not describe the underlying phenomena well.
In either case, we would not want to claim precise knowledge of the 
fitted coefficients for an economic growth model. 
The resampling technique that we employed
provides looser coefficient distributions when either of the above
is true. After discussing the fitting technique for each 
growth model below, we show scatter plots of fitted coefficients,
thereby providing visual indication of the uncertainty 
of the coefficients for each model.

@@

\noindent with 
$\alpha = \min(a,b)$, 
$\beta = |b-a|$, and 
$\gamma = 1 - \max(a,b)$.
The reparameterization ensures that $\alpha + \beta + \gamma = 1$,
thereby providing constant returns to scale.
The \texttt{R} \citep{R} function \texttt{nls} was used to estimate 
values of $\lambda$, $a$, and $b$.
If the fitted value for $a$ ($b$) fell outside $[0,1]$, it was 
set equal to its boundary value and $\lambda$ and $b$ ($a$) 
were re-estimated.

@@

>>>>

Under that assumption, it would not be surprising if the residual from 
any particular year had been observed in a different year. 
By applying randomly selected residuals [$\mathrm{abs}(\mathrm{rand}(r_t))$]
with randomly selected signs [$\mathrm{rand}(-1, 1)$]
to model predictions, we create a 
preliminary resampled GDP time series ($y_{prelim,resample,t}$):

\begin{equation} \label{eq:resample_y_prelim}
  y_{prelim,resample,t} = y_{pred,t} + \mathrm{rand}(-1, 1) \cdot \mathrm{abs}(\mathrm{rand}(r_t)).
\end{equation}

\noindent{}The preliminary resample time series
can have non-unity first-year GDP ($y_{prelim,resample,0} \ne 1$).
If we had actually obtained the resampled GDP data series,
we would have indexed to a different first-year GDP value, 
such that \emph{every} resampled GDP time series starts with $y_{resample,0} = 1$.
We use

\begin{equation} \label{eq:resample_y}
  y_{resample,t} = \frac{y_{prelim,resample,t}}{y_{prelim,resample,0}}
\end{equation}

\noindent{}to obtain a resampled indexed GDP time series that can be fitted using the techniques
described below. 
We repeated the resampling process \Sexpr{numResamples()} times for each
combination of growth model and country, both with and without energy when appropriate.

@@

It is important to note that large residuals can arise from either
(a) poor quality historical time series data ($y_{act}$) or 
(b) a mathematical model that does not describe the underlying phenomena well.
In both cases, we would not want to claim precise knowledge of the 
fitted coefficients for an economic growth model. 
So, it is appropriate that the resampling technique that we employed
provides looser coefficient distributions when either (a) or (b) above
is true. 
After discussing the fitting technique for each 
growth model below, we show scatter plots of fitted coefficients,
thereby providing visual indication of the uncertainty 
of the coefficients for each model.

@@

\noindent with 
$\alpha = \min(a,b)$, 
$\beta = |b-a|$, and 
$\gamma = 1 - \max(a,b)$.
The reparameterization ensures that $\alpha + \beta + \gamma = 1$,
thereby providing constant returns to scale.
The \texttt{R} \citep{R} function \texttt{nls} was used to estimate 
values of $\lambda$, $a$, and $b$.
If the fitted value for $a$ ($b$) fell outside $[0,1]$, it was 
set equal to its boundary value and both $\lambda$ and $b$ ($a$) 
were re-estimated.

@@

**** Replace with a spaghetti graph? Countries in rows. Without energy and with energy in columns. ****
